---
title: "R is better than MATLAB"
author: "Alex Holcombe"
date: "8/24/2017"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE,  comment = NA)
# http://htmlpreview.github.io/?https://github.com/alexholcombe/MixtureModelRSVP/blob/master/compareMATLABtoR.html
```

Import data

```{r importData, echo=FALSE}
#rm(list=ls())

mixModelingPath<- file.path("mixtureModeling")

#Import MATLAB fits
MATLABmixtureModelOutputPath<-"~/Google\ Drive/Backwards\ paper/secondPaper/E1/Data/MixModData"
importedToRbyChris <- "allParams.RData"
MATLABmixtureModelOutput<- file.path( MATLABmixtureModelOutputPath, importedToRbyChris )
  
#Import raw data
rawDataPath<- file.path("data/")
#Experiment was administered by MATLAB
#.mat file been preprocessed into melted long dataframe by backwarsLtrsLoadRawData.R
data<- readRDS( file.path(rawDataPath, "alexImportBackwardsPaper2E1.Rdata") ) 

#tidy data
library(dplyr)
df<- data
#It seems that to work with dplyr, can't have array field like letterSeq
df$letterSeq<- NULL


numItemsInStream<- length( data$letterSeq[1,] )  
minSPE<- -17; maxSPE<- 17

df<-df %>% rename(stream = target, orientation = condition )
df <- df %>% mutate( stream =ifelse(stream==1, "Left","Right") )
#mutate condition to Orientation
df <- df %>% mutate( orientation =ifelse(orientation==1, "Canonical","Inverted") )

#Load in MATLAB parameter estimates from Chris Bush run of MATLAB
load(MATLABmixtureModelOutput, verbose=FALSE)
#join into single long dataframe
estimates_MATLAB<- merge(efficacy.long,latency.long)
estimates_MATLAB<- merge(estimates_MATLAB,precision.long)
estimates_MATLAB<- data.frame(estimates_MATLAB)
names(estimates_MATLAB) <- tolower(names(estimates_MATLAB)) #lower case the column names
estimates_MATLAB<-estimates_MATLAB %>%
              mutate(efficacy=efficacy/100,latency=latency/100,precision=precision/100)


```

```{r calcLikelihoodForMATLABparams, echo=FALSE}
  #Want to add likelihood for MATLAB estimates to dataframe dfMl. Then can plot it with ggplot
  dfM<-merge( df, estimates_MATLAB )
  #dfM<- dfM %>% 
   #       rename(efficacy = efficacy_M, latency= latency_M, precision = precision_M)
  condtnVariableNames <- c("subject","stream", "orientation") # 

  #Unfortunately need the actual data for exact calculation of likelihood,
  # because guessing distribution depends on targetSP for that particular condition
  #Calculate the likelihood of the data given the MATLAB estimates
  source( file.path(mixModelingPath,"likelihoodOneConditionGivenParams.R") ) #for calcFitDataframes

  dfMl<- dfM %>%    # dplyr::filter(subject=="AA") %>%
    group_by_at(.vars = condtnVariableNames) %>%  #.dots needed when you use variable
    #do ( print(.) )
    do( likelihoodOneConditionForDplyr(.,numItemsInStream)   )
  #Calculated the likelihood for each condition.
  #Merge back with original dataframes.
  dfMl<- merge(dfM,dfMl)
  #Now need to reduce down, eliminating SPE so only one row per condition
  estimates_MATLAB <- unique( dfMl %>% select(stream,orientation,subject,efficacy,latency,precision,val) )
  #estimates_MATLAB<- merge(estimates_MATLAB, likEachCondtn)

```

Fit mixture model with R if not already available.
```{r, echo=FALSE, message=FALSE}
source( file.path(mixModelingPath,"analyzeOneCondition.R") )
source( file.path(mixModelingPath,"parameterBounds.R") )

if (!exists("estimates")) {
  estimates<- df %>%    #filter(subject=="AA") %>%
    group_by_(.dots = condtnVariableNames) %>%  #.dots needed when you have a variable containing multiple factor names
    do(analyzeOneCondition(.,numItemsInStream,parameterBounds()))
  estimates<- estimates %>% rename(efficacy = p1, latency = p2, precision = p3)
}
#head(estimates)
```

Unfortunately the differences between MATLAB and R are not trivial. Not substantial either! but we'd prefer trivial.

```{r, echo=FALSE, message=FALSE}
#Compare MATLAB parameter estimates to R
estimates_MATLAB<- estimates_MATLAB %>% rename(efficacy_M = efficacy, latency_M = latency, precision_M = precision, val_M = val)
merged<- merge(estimates,estimates_MATLAB)  
merged<- merged %>% mutate( effDiff = efficacy-efficacy_M, latDiff= latency_M-latency, preDiff= precision_M-precision, negLogL_Diff = val - val_M ) 

#show differences columns only
diffs<- select(merged, ends_with("Diff"))
```

The mean discrepancy (mean of the absolute value of the difference of the parameter estimates) for
efficacy (effDiff), latency (latDiff), and precision (preDiff) is
```{r, echo=FALSE, message=FALSE}
#calc discrepancies
#Don't include negLogL because absolute value not interpretable
meanDiscrepancy<- diffs %>% select(-negLogL_Diff) 
meanDiscrepancy<- summarise_all(abs(meanDiscrepancy),mean)
round(meanDiscrepancy,3)
```

Now we report the bias difference:  the mean difference without taking the absolute value. Note that the negLogLikelihood for the MATLAB estimates is calculated by R, because Pat's code doesn't save the negative log likelihoods. Also in R we now use integration of the whole area under the bin, whereas Pat used the height of the density function in the center of the bin.
```{r, echo=FALSE, message=FALSE}
biasDiff<- summarise_all(diffs,mean)
round(biasDiff,3)
```
Negative numbers above mean that MATLAB gives slightly higher efficacy and trivially higher precision.
For likelihood, lower neg log likelihood is better, so the negative number means R did better.

Let's inspect histogram plots and fits. Note that in every case, the negative log likelihood is better (smaller) for R, showing that the R code provides a better fit than the MATLAB code.

```{r, echo=FALSE, message=FALSE, fig.height=300}
#want fig.height of 10 per subject

source( file.path(mixModelingPath,"calcCurvesDataframes.R") )

#create R curves
df<- df  %>% dplyr::filter(subject=="AA")  #dplyr::filter(subject <="AP") #dplyr::filter(subject <= "BD" & subject >="AP")
#df<- df %>% dplyr::filter(subject=="AA",stream=="Right", orientation=="Canonical")
source( file.path(mixModelingPath,"histogramPlotting.R") ) #for calcFitDataframes
source( file.path(mixModelingPath,"theme_apa.R") ) #for calcFitDataframes

#Add R parameter estimates to dataframe
df<- merge(df,estimates) 

curvesR<- df %>% group_by_at(.vars = condtnVariableNames) %>% 
  do(calcCurvesDataframes(.,minSPE,maxSPE,numItemsInStream))

#Calc numObservations to each condition. This is needed only for scaling the fine-grained Gaussian
#Calc the number of observations for each condition, because gaussianScaledforData needs to know.
dfGroups<- df %>% group_by_at(.vars = condtnVariableNames) %>% summarise(nPerCond = n())
#add nPerCond back to parameter estimates
estimates<- merge(estimates,dfGroups)
grain<-.05
gaussianFineR<- estimates %>% group_by_at(.vars = condtnVariableNames) %>% do(
  gaussianScaledFromDataframe(.,minSPE,maxSPE,grain) )


#create MATLAB curves
#need to rename parameters so that calcCurvesDataframes can find them
estimates_M<- estimates_MATLAB %>% rename(efficacy=efficacy_M,latency=latency_M,precision=precision_M)
#create temporary dataframe with data plus MATLAB estimates to generate curves
dataStrippedOfEstimates<- select(df,-efficacy,-latency,-precision, -val)
dM<-merge( dataStrippedOfEstimates, estimates_M )
dM<- dM %>% rename( val = val_M ) #has to be called val so calcFitDataframes uses it rather than using R to calculate its own
curvesMATLAB<- dM %>% group_by_at(.vars = condtnVariableNames) %>% 
            do(calcCurvesDataframes(.,minSPE,maxSPE,numItemsInStream))

#Calculate fine-grained Gaussian for MATLAB
#Calc the number of observations for each condition, because gaussianScaledforData needs to know.
dNum<- dataStrippedOfEstimates %>% group_by_at(.vars = condtnVariableNames) %>% summarise(nPerCond = n())
#add nPerCond to MATLAB estimates, which will then be merged with raw data
estimates_M<- merge(estimates_M,dNum)
#merge MATLAB estimates with raw data
dM<-merge( dataStrippedOfEstimates, estimates_M )
grain<-.1
gaussianFineM<- estimates_M %>% group_by_at(.vars = condtnVariableNames) %>% do(
  gaussianScaledFromDataframe(.,minSPE,maxSPE,grain) )



#Concatenate R and MATLAB into single dataframe
curvesR$lang<-"R"; curvesMATLAB$lang<-"MATLAB"
curves_R_MATLAB <- rbind( data.frame(curvesR), data.frame(curvesMATLAB) )
#To facilitate plotting,
#make dataframe containing the raw data twice, one for R plots column, one for MATLAB plots column (data identical)
dfBoth<-rbind(   df %>% mutate(lang="R"), df %>% mutate(lang="MATLAB")  )

#Concatenate fine-grained gaussian for R and MATLAB into single dataframe, so can plot both with ggplot
gaussianFineR$lang<-"R"; gaussianFineM$lang<-"MATLAB"; 
#Get the col names to match by only using a few columns
gaussianFineR<- gaussianFineR %>% select(-val,-warnings)
gaussianFineM<- gaussianFineM %>% select(-val_M)
gaussFine_R_M <- rbind( data.frame(gaussianFineR), data.frame(gaussianFineM) )


g=ggplot(dfBoth, aes(x=SPE)) + facet_grid(subject+stream+orientation~lang)
g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
g<-g +theme_apa() #+theme(panel.grid.minor=element_blank(),panel.grid.major=element_blank())# hide all gridlines.
#g<-g+ theme(line=element_blank(), panel.border = element_blank())
sz=.8
g<-g+ geom_point(data=curves_R_MATLAB,aes(x=x,y=combinedFitFreq),color="chartreuse3",size=sz*2.5)
g<-g+ geom_line(data=curves_R_MATLAB,aes(x=x,y=guessingFreq),color="yellow",size=sz)
g<-g+ geom_line(data=curves_R_MATLAB,aes(x=x,y=gaussianFreq),color="lightblue",size=sz)


g<-g + geom_line(data=gaussFine_R_M,aes(x=x,y=gaussianFreq),color="darkblue",size=1.2)
#Plot the underlying Gaussian too, not just the discretized Gaussian
show(g)



numGroups<- length(table(dfBoth$orientation,dfBoth$subject,dfBoth$stream,dfBoth$lang))
fontSz = 80/numGroups
g<-g + geom_text(data=curves_R_MATLAB,aes(x=-9,y=32, label = paste("-logLik==", round(val,1), sep = "")),  parse = TRUE,size=fontSz) +
  geom_text(data=curves_R_MATLAB,aes(x=-7,y=28, label = paste("plain(e)==", round(efficacy,2), sep = "")),  parse = TRUE,size=fontSz) +
  geom_text(data=curves_R_MATLAB,aes(x=-7,y=25, label = paste("mu==", round(latency,2), sep = "")),  parse = TRUE,size=fontSz)+
  geom_text(data=curves_R_MATLAB,aes(x=-7,y=22, label = paste("sigma==", round(precision,2), sep = "")),  parse = TRUE,size=fontSz)
show(g)
```

